---
title: "Pharmacy First Consultations in OpenSAFELY-TPP"
output:
  html_document:
    toc: true
toc-title: "Overview"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(readr)
library(gt)
library(patchwork)
library(scales)
library(ggrepel)
```

```{r load-data, message=FALSE, warning=FALSE}
# Load functions
source(here("lib", "functions", "tidy_measures.R"))
source(here("lib", "functions", "plot_measures.R"))

# Load validation data:
# - df_bsa_medication_validation: date, pharmacy_advanced_service, bnf_paragraph, count
# - df_bsa_consultation_validation: date, consultation_type, source, count_method, count
source(here("lib", "functions", "load_validation_data.R"))

# Load opensafely ouputs:
# - df_measures: measure, interval_start, interval_end, ratio numerator, denominator, age_band, sex,imd, region, ethnicity
# - df_descriptive_stats: measure, interval_start, interval_end, ratio numerator, denominator
# - df_pfmed: measure, interval_start, interval_end, ratio, numerator, denominator, dmd_code
# - df_condition_provider: measure, interval_start, interval_end, ratio, numerator, denominator, pf_status, imd
source(here("lib", "functions", "load_opensafely_outputs.R"))
```

# Background

- COPY FROM PAPER (SHORT VERSION)

# Methods

- COPY FROM PAPER (SHORT VERSION)

# Results

## Count of Pharmacy First Consultations

### Total

```{r plot-pf-consultations, message=FALSE, warning=FALSE, fig.height=8, fig.width=10, fig.cap="Figure 1. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP."}
# Create figure for total count of Pharmacy First consultations for each code (3 codes)
df_pf_consultations <- df_measures |>
  filter(measure_desc == "clinical_service") |>
  filter(is.na(group_by)) |>
  select(measure, interval_start, numerator) |>
  mutate(measure = factor(measure,
    levels = c(
      "Consultation Service",
      "Pharmacy First Consultation",
      "Community Pharmacy First Service"
    ),
    labels = c(
      "CP Consultation Service for minor illness (1577041000000109)",
      "Pharmacy First service (983341000000102)",
      "CP Pharmacy First Service (2129921000000100)"
    )
  )) |>
  group_by(interval_start) |>
  mutate(
    pf_consultation_total = sum(numerator, na.rm = TRUE),
    data_desc = "Pharmacy First Consultation"
  ) |>
  rename(pf_code = measure) |>
  filter(interval_start >= "2024-02-01") |>
  ungroup() |>
  pivot_longer(
    cols = c(numerator, pf_consultation_total),
    names_to = "measure"
  ) |>
  mutate(
    measure = case_when(
      measure == "pf_consultation_total" ~ "Pharmacy First Service Total",
      measure == "numerator" ~ pf_code
    ),
    data_desc = case_when(
      measure == "Pharmacy First Service Total" ~ "total",
      measure %in% c(
        "CP Consultation Service for minor illness (1577041000000109)",
        "Pharmacy First service (983341000000102)",
        "CP Pharmacy First Service (2129921000000100)"
      ) ~ "breakdown"
    ),
    measure = factor(measure,
      levels = c(
        "Pharmacy First Service Total",
        "CP Consultation Service for minor illness (1577041000000109)",
        "Pharmacy First service (983341000000102)",
        "CP Pharmacy First Service (2129921000000100)"
      )
    ),
    data_desc = factor(data_desc,
      levels = c("total", "breakdown"),
      labels = c("Total", "Breakdown")
    )
  ) |>
  select(measure, interval_start, data_desc, value)

plot_pf_consultations <- df_pf_consultations |>
  ggplot(
    aes(
      x = interval_start,
      y = value,
      colour = measure,
      shape = measure,
      fill = measure
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(alpha = .2) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  guides(
    color = guide_legend(nrow = 2),
    shape = guide_legend(nrow = 2)
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = NULL,
    colour = NULL,
    shape = NULL,
    fill = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14)
  ) +
  facet_wrap(
    ~data_desc,
    ncol = 1
  ) +
  scale_colour_viridis_d(option = "plasma", end = .75) +
  scale_y_continuous(
    labels = scales::number,
    breaks = c(seq(0, 60000, by = 10000))
  )

plot_pf_consultations
```

### Demographic breakdowns

#### Sex

```{r plot-pf-consultations-by-sex, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="Figure 2. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by sex."}
df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Sex") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = sex,
    guide_nrow = 1,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  )
```

#### Age band

```{r plot-pf-consultations-by-age-band, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="Figure 3. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by age bands."}
unique_categories <- sort(unique(df_measures$age_band))
manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 1)
shape_mapping <- setNames(manual_ggplot2_shapes[1:length(unique_categories)], unique_categories)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Age band") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = age_band,
    guide_nrow = 1,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping)
```

#### Index of Multiple Deprivation (IMD)

```{r plot-pf-consultations-by-imd, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="Figure 4. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by IMD."}
unique_categories <- sort(unique(df_measures$imd))

manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 1)
shape_mapping <- setNames(
  manual_ggplot2_shapes[1:length(unique_categories)],
  unique_categories
)

# Get one more colour than I need from viridis so I avoid using yellow in the plot
# paletteer_c("viridis::viridis", n = 7)
manual_ggplot2_colours <- c(
  "#440154FF",
  "#443A83FF",
  "#31688EFF",
  "#21908CFF",
  "#35B779FF",
  "#8FD744FF",
  "#FDE725FF"
)
colour_mapping <- setNames(
  c(manual_ggplot2_colours[1:(length(unique_categories) - 1)], "gray40"),
  unique_categories
)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "IMD") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = imd,
    guide_nrow = 1,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping) +
  scale_colour_manual(values = colour_mapping) +
  scale_fill_manual(values = colour_mapping)
```

#### Region

```{r plot-pf-consultations-by-region, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="Figure 5. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by region."}
unique_categories <- sort(unique(df_measures$region))

manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 8, 4, 10, 7, 1)
shape_mapping <- setNames(
  manual_ggplot2_shapes[1:length(unique_categories)],
  unique_categories
)

# Get one more colour than I need from viridis so I avoid using yellow in the plot
# paletteer::paletteer_c("viridis::viridis", n = 11)
manual_ggplot2_colours <- c(
  "#440154FF",
  "#482576FF",
  "#414487FF",
  "#35608DFF",
  "#2A788EFF",
  "#21908CFF",
  "#22A884FF",
  "#43BF71FF",
  "#7AD151FF",
  "#BBDF27FF",
  "#FDE725FF"
)

colour_mapping <- setNames(
  c(manual_ggplot2_colours[1:(length(unique_categories) - 1)], "gray40"),
  unique_categories
)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Region") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = region,
    guide_nrow = 2,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping) +
  scale_colour_manual(values = colour_mapping) +
  scale_fill_manual(values = colour_mapping)
```

#### Ethnicity

```{r plot-pf-consultations-by-ethnicity, message=FALSE, warning=FALSE, fig.height=15, fig.width=10, fig.cap="Figure 5. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by region."}
unique_categories <- sort(unique(df_measures$ethnicity))

manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 8, 4, 10, 7, 1)
shape_mapping <- setNames(
  manual_ggplot2_shapes[1:length(unique_categories)],
  unique_categories
)

# Get one more colour than I need from viridis so I avoid using yellow in the plot
# paletteer::paletteer_c("viridis::viridis", n = 6)
manual_ggplot2_colours <- c(
  "#440154FF",
  "#414487FF",
  "#2A788EFF",
  "#22A884FF",
  "#7AD151FF",
  "#FDE725FF"
)

colour_mapping <- setNames(
  c(manual_ggplot2_colours[1:(length(unique_categories) - 1)], "gray40"),
  unique_categories
)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Ethnicity") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = ethnicity,
    guide_nrow = 2,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping) +
  scale_colour_manual(values = colour_mapping) +
  scale_fill_manual(values = colour_mapping) +
  facet_wrap(~ethnicity, ncol = 1, scales = "free_y")
```

## Linkage of Pharmacy First Consultations

```{r plot-pf-med-condition-linkage, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="Figure 6. Monthly breakdown of Pharmacy First consultation codes linked to a Pharmacy First medication, clinical condition, or both in OpenSAFELY-TPP."}
df_pf_descriptive_stats <- df_descriptive_stats %>%
  filter(measure %in% c("pfmed_with_pfid", "pfcondition_with_pfid", "pfmed_and_pfcondition_with_pfid")) %>%
  mutate(
    measure = factor(measure,
      levels = c(
        "pfmed_with_pfid",
        "pfcondition_with_pfid",
        "pfmed_and_pfcondition_with_pfid"
      ),
      labels = c(
        "Medication",
        "Clinical condition",
        "Both"
      )
    )
  )

df_pf_descriptive_stats <- df_pf_descriptive_stats %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) %>%
  group_by(interval_start) %>%
  left_join(df_pf_consultations_total,
    by = c("interval_start")
  ) %>%
  arrange(desc(measure), .by_group = TRUE) %>%
  mutate(
    ratio = numerator / pf_consultation_total,
    cumulative_ratio = case_when(interval_start == max(df_pf_descriptive_stats$interval_start) ~ cumsum(ratio), TRUE ~ NA)
  )

plot_pf_med_condition_linkage <- df_pf_descriptive_stats |>
  ggplot(
    aes(
      x = interval_start,
      y = ratio,
      fill = measure
    )
  ) +
  geom_area(
    alpha = 0.85,
    linewidth = .5,
    colour = "white"
  ) +
  geom_label(
    aes(
      y = cumulative_ratio,
      label = ifelse(is.na(ratio), "", scales::percent(ratio, accuracy = .1)),
    ),
    vjust = -0.05,
    size = 3.5,
    show.legend = FALSE,
    color = "black",
    fill = "white"
  ) +
  scale_fill_viridis_d() +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Pharmacy First consultations linked to: "
  ) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14)
  )

plot_pf_med_condition_linkage
```

### Pharmacy First Clinical Pathways

- Add existing breakdown figures for: Age band, Ethnicity, IMD

### Pharmacy First Medications

- Add existing table from paper once updated with released outputs

## Comparison with NHS BSA data

### Pharmacy First Consultations

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="Figure 7. Monthly counts and ratio of patients with a Pharmacy First consultation code in OpenSAFELY-TPP and the NHS BSA comparison dataset."}
# Ideally I would like to load this at the top of this script
# But this file uses datasets created further down so it easiest to load it here
df_pf_consultations_total <- df_pf_consultations |>
  filter(data_desc == "Total") |>
  mutate(
    data_desc = "Pharmacy First Consultation"
  ) |>
  select(interval_start, pf_consultation_total = value, data_desc) |>
  unique()

source(here("lib", "functions", "combine_os_nhsbsa_validation_data.R"))

# Pharmacy First Consultations
# Instead of only using the count of PF consultations in OS with a clinical
# condition we are using all PF consultations with a consultation id
df_pf_consultation_validatio_plot <- pf_validation |>
  filter(
    (data_source == "opensafely" &
      data_desc == "pf_consultation" &
      count_desc == "consultation_id") |
      (data_source == "nhs_bsa" &
        data_desc == "pf_consultation" &
        count_desc == "consultation_type")
  )

# Pivot wider to calculate ratio
df_pf_consultation_validatio_plot <- df_pf_consultation_validatio_plot |>
  select(date, data_source, count_method, count_total) |>
  pivot_wider(names_from = c(data_source, count_method), values_from = count_total) |>
  mutate(ratio = opensafely_opensafely_tpp / nhs_bsa_count_100pct)

# Select colours for groups
# Group 1:
colour_group_1 <- "#CF4446FF"
# Group 1: #1B0C42FFpaletteer_c("viridis::inferno", n = 10)[2]
colour_group_2 <- "#1B0C42FF"

# Create consultation validation plot
pf_validatio_consultation_plot <- df_pf_consultation_validatio_plot |>
  ggplot(aes(x = date)) +
  geom_point(
    aes(
      y = nhs_bsa_count_100pct,
      colour = "NHS BSA",
      shape = "NHS BSA"
    ),
    size = 2.6
  ) +
  geom_line(
    aes(y = nhs_bsa_count_100pct, colour = "NHS BSA"),
    alpha = .3,
    linewidth = .5
  ) +
  geom_point(
    aes(
      y = opensafely_opensafely_tpp,
      colour = "OpenSAFELY-TPP",
      shape = "OpenSAFELY-TPP"
    ),
    size = 2.6
  ) +
  geom_line(
    aes(
      y = opensafely_opensafely_tpp,
      colour = "OpenSAFELY-TPP"
    ),
    alpha = .3,
    linewidth = .5
  ) +
  geom_label(
    aes(
      y = opensafely_opensafely_tpp,
      label = scales::percent(ratio, accuracy = .1),
    ),
    colour = colour_group_2,
    vjust = -.5,
    size = 3.5
  ) +
  scale_y_continuous(labels = scales::number) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  scale_colour_manual(
    values = c(
      "NHS BSA" = colour_group_1,
      "OpenSAFELY-TPP" = colour_group_2
    ),
    name = NULL
  ) +
  scale_shape_manual(
    values = c(
      "NHS BSA" = 15,
      "OpenSAFELY-TPP" = 16
    ),
    name = NULL
  ) +
  labs(
    x = NULL,
    y = NULL,
    colour = NULL,
    shape = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14)
  )

pf_validatio_consultation_plot
```


### Pharmacy First Medications
