---
title: "Pharmacy First"
output:
  html_document:
    toc: true
    toc_depth: 4
  pdf_document: default
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: references.bib
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(readr)
library(gt)
library(patchwork)
```

```{r load-data, message=FALSE, warning=FALSE}
# Load functions
source(here("lib", "functions", "tidy_measures.R"))
source(here("lib", "functions", "plot_measures.R"))

# Load validation data:
# - df_bsa_medication_validation: date, pharmacy_advanced_service, bnf_paragraph, count
# - df_bsa_consultation_validation: date, consultation_type, source, count_method, count
source(here("lib", "functions", "load_validation_data.R"))

# Load opensafely ouputs:
# - df_measures: measure, interval_start, interval_end, ratio numerator, denominator, age_band, sex,imd, region, ethnicity
# - df_descriptive_stats: measure, interval_start, interval_end, ratio numerator, denominator
# - df_pfmed: measure, interval_start, interval_end, ratio, numerator, denominator, dmd_code
# - df_condition_provider: measure, interval_start, interval_end, ratio, numerator, denominator, pf_status, imd
source(here("lib", "functions", "load_opensafely_outputs.R"))
```

# Background

Add background here.

# Methods

## Data source

Primary care records managed by the GP software provider TPP were accessed through OpenSAFELY (https://opensafely.org).
OpenSAFELY-TPP covers approximately 40% of the population of England, for a detailed description of the representativeness see @Andrews2022.

Between 1st November 2023 and 1st September 2024, we identified individuals aged between 0 and 120 years who were registered at a TPP practice each month across the study period. 
From this cohort, we counted the number of patients with at least one SNOMED CT code indicating a consultation using the Pharmacy First Service. 
Patients with more than one instance of the same clinical code within a month were counted only once.

Our study relied on the correct entry of relevant Pharmacy First codes in patients' GP records. 
The Pharmacy First service uses [GP Connect - Update Record](https://digital.nhs.uk/services/gp-connect/gp-connect-in-your-organisation/gp-connect-update-record) to update a patient's GP record with consultation information from community pharmacies. 
Following the launch of the Pharmacy First service, there has been a [Gradual roll-out of GP Connect - Update Record](https://cpe.org.uk/our-news/gp-connect-update-record-rollout-and-flow-of-information/) across approved community pharmacy IT system suppliers.

### Population

The eligible patient population for each clinical condition associated with Pharmacy First is detailed in the table below.
Currently, we have not yet applied any of the inclusion or exlusion citeria to restrict the population used in this report.
This will initially help us to understand the underlying data.

```{r echo=FALSE, message=FALSE}
# Create clinical pathways dataframe
clinical_pathways_table <- data.frame(
  Condition = c(
    "Uncomplicated Urinary Tract Infection",
    "Shingles",
    "Impetigo",
    "Infected Insect Bites",
    "Acute Sore Throat",
    "Acute Sinusitis",
    "Acute Otitis Media"
  ),
  Age = c(
    "16 to 64 years",
    "18 years and over",
    "1 year and over",
    "1 year and over",
    "5 years and over",
    "12 years and over",
    "1 to 17 years"
  ),
  Sex = c(
    "Female",
    "Any",
    "Any",
    "Any",
    "Any",
    "Any",
    "Any"
  ),
  Exclusions = c(
    "Pregnant individuals, urinary catheter, recurrent UTI (2 episodes in last 6 months, or 3 episodes in last 12 months)",
    "Pregnant individuals",
    "Bullous impetigo, recurrent impetigo (2 or more episodes in the same year), pregnant individuals under 16 years",
    "Pregnant individuals under 16 years",
    "Pregnant individuals under 16 years",
    "Immunosuppressed individuals, chronic sinusitis (symptoms lasting more than 12 weeks), pregnant individuals under 16 years",
    "Recurrent acute otitis media (3 or more episodes in 6 months or four or more episodes in 12 months), pregnant individuals under 16 years"
  )
)

# Create clinical pathways table
clinical_pathways_table %>%
  gt() %>%
  tab_header(
    title = "Table 1. Pharmacy First population criteria"
    # subtitle = "Inclusion and exclusion criteria for clinical pathway/conditions"
  ) %>%
  cols_label(
    Condition = "Condition",
    Age = "Age Range",
    Sex = "Sex",
    Exclusions = "Exclusions"
  ) %>%
  tab_options(
    table.font.size = "medium",
    heading.title.font.size = "large",
    heading.subtitle.font.size = "small"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  )
```

### Codelists

We used the following codelists to identify Pharmacy First consultations, conditions, and demographic breakdowns.

#### Pharmacy First consultation codes

The following two SNOMED codes were used to identify Pharmacy First consultations.
For clarity, we combined these codes for the presentation of the results.

```{r echo=FALSE}
# Create pharmacy first service codes dataframe
pharmacy_first_table <- data.frame(
  codelist = c(
    "Community Pharmacist (CP) Consultation Service for minor illness (procedure)",
    "Pharmacy First service (qualifier value)"
  ),
  code = c(
    "1577041000000109",
    "983341000000102"
  )
)

# Create pharmacy first service codes table
pharmacy_first_table %>%
  gt() %>%
  tab_header(
    title = "Table 2. Pharmacy First consultation codes",
    # subtitle = "Codelist descriptions and their respective SNOMED codes"
  ) %>%
  cols_label(
    codelist = md("**Codelist Description**"),
    code = md("**SNOMED Code**")
  ) %>%
  tab_options(
    table.font.size = "medium",
    heading.title.font.size = "large",
    heading.subtitle.font.size = "small"
  )
```

#### Pharmacy First condition codes

To categorise clinical events related to Pharmacy First services used the Pharmacy First [Clinical Pathways Codelist](https://www.opencodelists.org/codelist/opensafely/pharmacy-first-clinical-pathway-conditions/7ec97762/#full-list).

```{r echo=FALSE, message=FALSE}
clinical_codes_table <- data.frame(
  condition = c(
    "Acute otitis media",
    "Herpes zoster",
    "Acute sinusitis",
    "Impetigo",
    "Infected insect bite",
    "Acute pharyngitis",
    "Uncomplicated urinary tract infection"
  ),
  code = c(
    "3110003",
    "4740000",
    "15805002",
    "48277006",
    "262550002",
    "363746003",
    "1090711000000102"
  )
)

clinical_codes_table %>%
  gt() %>%
  tab_header(
    title = "Table 3. Pharmacy First condition codes"
    # subtitle = "Clinical conditions and their corresponding SNOMED codes"
  ) %>%
  cols_label(
    condition = md("**Clinical Condition**"),
    code = md("**SNOMED Code**")
  ) %>%
  tab_options(
    table.font.size = "medium",
    heading.title.font.size = "large",
    heading.subtitle.font.size = "small"
  )
```

#### Pregnancy Codelist

The [Pregnancy Codelist](https://www.opencodelists.org/codelist/nhsd-primary-care-domain-refsets/preg_cod/20200812/#full-list) was used to identify patients who were pregnant during each month.

#### Ethnicity Codelist

We used the [Ethnicity Codelist](https://www.opencodelists.org/codelist/opensafely/ethnicity-snomed-0removed/2e641f61/) identify ethnicity in Electronic Health Records.
To ensure comprehensive ethnicity data, we supplemented missing ethnicity values with data from the Secondary Uses Service (SUS).

# Results

### Total population

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=10}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(is.na(group_by)) |>
  select(measure, interval_start, numerator) |>
  mutate(measure = factor(measure,
    levels = c("Consultation Service", "Pharmacy First Consultation"),
    labels = c(
      "Consultation Service for minor illness (1577041000000109)",
      "Pharmacy First service (983341000000102)"
    )
  ))

fig_pf_individual_consultations_count <- df_measures_selected |>
  select(measure, interval_start, numerator) |>
  ggplot(aes(
    x = interval_start,
    y = numerator,
    colour = measure,
    shape = measure,
  )) +
  geom_point(size = 2) +
  geom_line(alpha = .3) +
  labs(
    title = NULL,
    x = NULL,
    y = "Total count",
    colour = NULL,
    shape = NULL
  ) +
  scale_y_continuous(
    labels = scales::label_number(),
  ) +
  theme(legend.position = "bottom") +
  guides(
    colour = guide_legend(ncol = 2),
    shape = guide_legend(ncol = 2)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  geom_vline(
    xintercept = lubridate::as_date(c(
      "2024-01-31"
    )),
    linetype = "dotted",
    colour = "orange",
    size = .7
  ) +
  scale_colour_viridis_d(end = .75) +
  theme(
    text = element_text(size = 14)
  )


ggsave(
  here("released_output", "results", "figures", "fig_pf_individual_consultations_count.png"),
  fig_pf_individual_consultations_count,
  width = 10, height = 6
)

fig_pf_individual_consultations_count

fig_pf_grouped_consultations_count <- df_measures_selected |>
  group_by(interval_start) |>
  mutate(
    pf_consultation_total = sum(numerator, na.rm = TRUE),
    data_desc = "Pharmacy First Consultation"
  ) |>
  ggplot(aes(
    x = interval_start,
    y = pf_consultation_total,
    colour = data_desc,
    shape = data_desc,
  )) +
  geom_point(size = 2) +
  geom_line(alpha = .3) +
  labs(
    title = NULL,
    x = NULL,
    y = "Total count",
    colour = NULL,
    shape = NULL
  ) +
  scale_y_continuous(
    labels = scales::label_number(),
  ) +
  theme(legend.position = "bottom") +
  guides(
    colour = guide_legend(ncol = 2),
    shape = guide_legend(ncol = 2)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  geom_vline(
    xintercept = lubridate::as_date(c(
      "2024-01-31"
    )),
    linetype = "dotted",
    colour = "orange",
    size = .7
  ) +
  scale_colour_viridis_d(end = .75) +
  theme(
    text = element_text(size = 14)
  )


ggsave(
  here("released_output", "results", "figures", "fig_pf_grouped_consultations_count.png"),
  fig_pf_grouped_consultations_count,
  width = 10, height = 6
)
```

```{r, message=FALSE, warning=FALSE, fig.height=10, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(is.na(group_by))

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions",
)
```

### Breakdown by age

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)
```

### Breakdown by sex

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = sex_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = sex_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = sex_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = sex_palette)
```

### Breakdown by IMD

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)
```

### Breakdown by region

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = region_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = region_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = region_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = region_palette)
```

### Breakdown by ethnicity

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = ethnicity_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = ethnicity_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = ethnicity_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = ethnicity_palette)
```
```{r, message=FALSE, warning=FALSE, echo = FALSE}
# OpenSAFELY data for clinical conditions into a tidy df
df_opensafely_validation <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  # filter(interval_start >= as.Date("2024-02-01") & interval_start <= as.Date("2024-07-30")) %>%
  filter(is.na(group_by)) %>%
  select(date = interval_start, consultation_type = measure, count = numerator) %>%
  mutate(
    source = "opensafely",
    count_method = "opensafely_tpp"
  ) |>
  filter(date >= "2024-01-01") %>%
  relocate(date, consultation_type, source, count_method, count)

# Combining rows from OS and BSA validation dataframes
df_validation_condition_counts <- bind_rows(df_opensafely_validation, df_bsa_consultation_validation)

# Line graph comparing clinical condition counts of BSA and OS data
fig_validation_condition_count <- df_validation_condition_counts %>%
  filter(count_method %in% c("opensafely_tpp", "count_40pct")) %>%
  mutate(source = factor(source,
    levels = c("opensafely", "nhs_bsa"),
    labels = c("OpenSAFELY-TPP", "NHS BSA (40%)")
  )) %>%
  ggplot(
    aes(
      x = date,
      y = count,
      shape = consultation_type,
      color = consultation_type,
      fill = consultation_type,
      group = consultation_type
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  facet_wrap(~source, scales = "free_y") +
  scale_x_date(
    labels = scales::label_date_short()
  ) +
  labs(x = NULL, y = "Count", colour = NULL, shape = NULL, fill = NULL) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_shape_manual(
    values = c(
      "Acute Sinusitis" = 15,
      "Infected Insect Bite" = 19,
      "UTI" = 4,
      "Acute Otitis Media" = 23,
      "Acute Pharyngitis" = 3,
      "Herpes Zoster" = 17,
      "Impetigo" = 8
    )
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_y_continuous(labels = scales::number)

# Another plot visualising the percentage
fig_validation_condition_pct <- df_validation_condition_counts %>%
  filter(count_method %in% c("opensafely_tpp", "count_40pct")) %>%
  pivot_wider(names_from = c(source, count_method), values_from = count) %>%
  mutate(source = "Percentage of NHS BSA (40%) in OpenSAFELY") %>%
  ggplot(
    aes(
      x = date,
      y = opensafely_opensafely_tpp / nhs_bsa_count_40pct,
      shape = consultation_type,
      color = consultation_type,
      fill = consultation_type,
      group = consultation_type
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  facet_wrap(~source, scales = "free_y") +
  scale_x_date(
    labels = scales::label_date_short()
  ) +
  labs(x = NULL, y = "Percent", colour = NULL, shape = NULL, fill = NULL) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_shape_manual(
    values = c(
      "Acute Sinusitis" = 15,
      "Infected Insect Bite" = 19,
      "UTI" = 4,
      "Acute Otitis Media" = 23,
      "Acute Pharyngitis" = 3,
      "Herpes Zoster" = 17,
      "Impetigo" = 8
    )
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_y_continuous(labels = scales::percent)

# Combining the plots with patchwork
fig_validation_condition_count_pct <- (fig_validation_condition_count + fig_validation_condition_pct) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect", widths = c(2, 1)) &
  theme(
    legend.position = "bottom",
    text = element_text(size = 15),
    strip.background = element_rect(size = 0),
    strip.text.x = element_text(size = 13, face = "bold")
  )

fig_validation_condition_count_pct

ggsave(
  here("released_output", "results", "figures", "fig_validation_condition_count_pct.png"),
  fig_validation_condition_count_pct,
  width = 15, height = 6
)
```

```{r, message=FALSE, warning=FALSE, echo = FALSE, fig.width=8}
# Line graph comparing clinical condition counts of BSA and OS data
fig_pf_descriptive_stats <- df_descriptive_stats %>%
  mutate(
    measure = factor(measure,
      levels = c("pf_with_pfmed", "pf_with_pfcondition", "pf_with_pfmed_and_pfcondition"),
      labels = c("PF Med", "PF Condition", "PF Med & PF Condition")
    )
  ) |>
  ggplot(aes(
    x = interval_start,
    y = ratio,
    colour = measure,
    shape = measure,
  )) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  labs(
    x = NULL,
    y = NULL,
    shape = "PF consultation linked to:",
    colour = "PF consultation linked to:"
  ) +
  scale_x_date(
    labels = scales::label_date_short(), breaks = "month"
  ) +
  scale_y_continuous(
    labels = scales::percent,
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_colour_brewer(palette = "Dark2")


fig_pf_descriptive_stats

ggsave(
  here("released_output", "results", "figures", "fig_pf_descriptive_stats.png"),
  fig_pf_descriptive_stats,
  width = 10, height = 6
)
```

```{r, message=FALSE, warning=FALSE, echo = FALSE}
# Validation of pharmacy first medication counts figure
# OS data - waiting on released output

df_bsa_medication_validation_sum <- df_bsa_medication_validation %>%
  group_by(date) %>%
  summarise(count = sum(count) * 0.4) %>%
  mutate(
    source = "nhs_bsa",
    count_method = "count_40pct"
  )
range(df_pfmed$interval_start)
df_opensafely_pfmed_sum <- df_pfmed %>%
  rename(date = interval_start) %>%
  group_by(date) %>%
  summarise(count = sum(numerator)) %>%
  mutate(
    source = "opensafely_tpp",
    count_method = "opensafely_tpp"
  )

df_validation_med_counts <- bind_rows(df_opensafely_pfmed_sum, df_bsa_medication_validation_sum) |>
  filter(date >= "2024-01-01" & date <= "2024-07-01")

fig_validation_med_count <- df_validation_med_counts |>
  mutate(
    source = factor(source, levels = c("opensafely_tpp", "nhs_bsa"), labels = c("OpenSAFELY-TPP", "NHS BSA")),
    count_method = factor(count_method, levels = c("opensafely_tpp", "count_40pct"), labels = c("OpenSAFELY-TPP", "NHS BSA (40%)"))
  ) |>
  ggplot(aes(
    x = date,
    y = count,
    colour = count_method,
    shape = count_method
  )) +
  geom_point(size = 2) +
  facet_wrap(~count_method, scales = "free_y") +
  geom_line(size = 0.5) +
  labs(
    x = NULL,
    y = "Count",
    colour = NULL,
    shape = NULL,
  ) +
  scale_x_date(
    labels = scales::label_date_short(), breaks = "month"
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_colour_brewer(palette = "Dark2")

# Another plot visualising the percentage
fig_validation_med_pct <- df_validation_med_counts %>%
  filter(count_method %in% c("opensafely_tpp", "count_40pct")) %>%
  pivot_wider(names_from = c(source, count_method), values_from = count) %>%
  mutate(source = "Percentage of NHS BSA (40%) in OpenSAFELY") %>%
  ggplot(
    aes(
      x = date,
      y = opensafely_tpp_opensafely_tpp / nhs_bsa_count_40pct,
      shape = source,
      color = source,
      fill = source,
      group = source
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  facet_wrap(~source, scales = "free_y") +
  scale_x_date(
    labels = scales::label_date_short(), breaks = "month"
  ) +
  labs(x = NULL, y = "Percent", colour = NULL, shape = NULL, fill = NULL) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_shape_manual(
    values = c("Percentage of NHS BSA (40%) in OpenSAFELY" = 15)
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_y_continuous(labels = scales::percent)

# Combining the plots with patchwork
fig_validation_medication_count_pct <- (fig_validation_med_count + fig_validation_med_pct) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect", widths = c(2, 1)) &
  theme(
    legend.position = "bottom",
    text = element_text(size = 15),
    strip.background = element_rect(size = 0),
    strip.text.x = element_text(size = 13, face = "bold")
  )

fig_validation_medication_count_pct

ggsave(
  here("released_output", "results", "figures", "fig_validation_medication_count_pct.png"),
  fig_validation_medication_count_pct,
  width = 15, height = 6
)
```
```{r, message=FALSE, warning=FALSE, echo = FALSE, fig.width=8}
# GP vs PF provider graph

df_condition_provider_grouped <- df_condition_provider %>%
  group_by(measure, interval_start, pf_status) %>%
  summarise(
    count = sum(numerator)
  ) %>%
  mutate(
    measure = factor(measure,
      levels = c(
        "count_acute_sinusitis_total",
        "count_infected_insect_bite_total",
        "count_uncomplicated_urinary_tract_infection_total",
        "count_acute_otitis_media_total",
        "count_acute_pharyngitis_total",
        "count_herpes_zoster_total",
        "count_impetigo_total"
      ),
      labels = c(
        "Acute Sinusitis",
        "Infected Insect Bite",
        "UTI",
        "Acute Otitis Media",
        "Acute Pharyngitis",
        "Herpes Zoster",
        "Impetigo"
      )
    ),
    pf_status = factor(pf_status,
      levels = c(TRUE, FALSE),
      labels = c("Linked to Pharmacy First consultation", "Not linked to Pharmacy First consultation")
    )
  )

fig_pf_condition_provider_count <- ggplot(
  df_condition_provider_grouped,
  aes(
    x = interval_start,
    y = count,
    colour = pf_status,
    shape = pf_status
  )
) +
  geom_point(size = 1.5) +
  geom_line(size = 0.5) +
  facet_wrap(~measure, scales = "free_y") +
  labs(
    x = NULL, y = "Count", color = NULL, shape = NULL
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    axis.title.x = element_blank()
  ) +
  scale_x_date(
    labels = scales::label_date_short()
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.75
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14),
    strip.background = element_rect(size = 0),
    # strip.text.x = element_text(size = 13, face = "bold")
  )

fig_pf_condition_provider_count

ggsave(
  here("released_output", "results", "figures", "fig_pf_condition_provider_count.png"),
  fig_pf_condition_provider_count,
  width = 13, height = 8
)
```

# References
