---
title: "Pharmacy First Consultations in OpenSAFELY-TPP"
output:
  html_document:
    toc: true
toc-title: "Overview"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "references.bib"
link-citations: true
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(readr)
library(gt)
library(patchwork)
library(scales)
library(ggrepel)
```

```{r load-data, message=FALSE, warning=FALSE}
# Load functions
source(here("lib", "functions", "tidy_measures.R"))
source(here("lib", "functions", "plot_measures.R"))
source(here("lib", "functions", "create_tables.R"))

# Load validation data:
# - df_bsa_medication_validation: date, pharmacy_advanced_service, bnf_paragraph, count
# - df_bsa_consultation_validation: date, consultation_type, source, count_method, count
source(here("lib", "functions", "load_validation_data.R"))

# Load opensafely ouputs:
# - df_measures: measure, interval_start, interval_end, ratio numerator, denominator, age_band, sex,imd, region, ethnicity
# - df_descriptive_stats: measure, interval_start, interval_end, ratio numerator, denominator
# - df_pfmed: measure, interval_start, interval_end, ratio, numerator, denominator, dmd_code
# - df_condition_provider: measure, interval_start, interval_end, ratio, numerator, denominator, pf_status, imd
source(here("lib", "functions", "load_opensafely_outputs.R"))

df_pf_consultations <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(is.na(group_by)) |>
  select(measure, interval_start, numerator) |>
  mutate(measure = factor(measure,
    levels = c("Consultation Service", "Pharmacy First Consultation", "Community Pharmacy First Service"),
    labels = c(
      "Consultation Service for minor illness (1577041000000109)",
      "Pharmacy First service (983341000000102)",
      "Community Pharmacy First Service (2129921000000100)"
    )
  ))

df_pf_consultations_total <- df_pf_consultations |>
  group_by(interval_start) |>
  mutate(
    pf_consultation_total = sum(numerator, na.rm = TRUE),
    data_desc = "Pharmacy First Consultation"
  ) %>%
  filter(interval_start >= "2024-02-01") %>%
  select(interval_start, pf_consultation_total, data_desc) |>
  ungroup() |>
  distinct()
```

# Background

The COVID-19 pandemic has had a significant impact on the access to and provision of NHS healthcare services.
In response to this, NHS England announced a recovery plan to "*make it easier and quicker for patients to get the help they need from primary care*" [@NHSEngland2023a].

A key element of the recovery plan is the Pharmacy First scheme [@NHSEngland2024], launched on 31st January 2024 which replaces and expands upon the Community Pharmacist Consultation Service (CPCS), which ran from 29th Oct 2019 [@NHSEngland2024a].
CPCS incorporated two services - Urgent Repeat Medicine Supply and NHS Referrals for Minor Illness - these services continue under Pharmacy First but there is also a new clinical pathways service [@NHSEngland2023].
This aspect allows community pharmacists to follow specified clinical pathways to assess for and manage seven common conditions (Sinusitis, Sore throat, Earache, Infected insect bite, Impetigo, Shingles, and Uncomplicated urinary tract infections in women), including supplying NHS funded prescription only medications where appropriate [@NHSEngland2024]. 

As part of efforts to improve and simplify information sharing between general practice and community pharmacy GP Connect: Update Record has been created, allowing registered community pharmacy professionals to send summaries in a structured format from Pharmacy First consultations, using pre-specified SNOMED CT codes, directly into general practice workflow [@NHSEngland2023b].
However, there has been a gradual roll out of Update Record and it has also been the target of GP collective action adding some uncertainty about how reliably data has flowed into general practice records following the launch of the Pharmacy First service, with reliance on manual transfer and recording of consultation summaries where Update Record was not available [@Colivicchi2024; @CommunityPharmacyEngland2024].

The aims of this report are to:

1. describe the clinical data associated with Pharmacy First Clinical Pathways in GP records including demographic and regional subgroup descriptions of patients
2. describe the clinical conditions and medications being treated/supplied under Pharmacy First Clinical Pathways and recorded in GP records,
3. compare volumes of Pharmacy First Clinical Pathways activity taking place in GP records accessible via OpenSAFELY-TPP against openly available sources of Pharmacy First Clinical Pathways data,

# Methods

## Data Sources
### OpenSAFELY-TPP

Primary care records managed by the GP software provider TPP were accessed through OpenSAFELY OpenSAFELY (https://www.opensafely.org/), a secure analytics platform for EHRs built during the COVID-19 pandemic.
OpenSAFELY operates within the data centres of EHR providers, allowing for analysis without transferring raw data outside the secure environment.
Data were accessed from OpenSAFELY-TPP, which covers approximately 40% of the English population, and includes pseudonymised, structured data 
The data used comprised coded diagnoses, medications and physiological parameters. Data from individuals who opted out of NHS data sharing with their GP surgery (Type 1 Opt-out) were excluded and free-text data were also not included.

### NHS BSA comparison data
To explore the accuracy and completeness of the data used in this study, we compared Pharmacy First consultation counts and trends observed in OpenSAFELY-TPP against equivalent metrics reported by NHS Business Services Authority (NHS BSA), which provides publically available records of dispensing activity and service usage.
We used the Pharmacy and appliance contractor dispensing dataset which includes monthly counts of Pharmacy First Clinical Pathways [@NHSBusinessServiceAuthority2025].

## Study Measures
### Pharmacy First consultations

The identification of Pharmacy First consultations relied on the presence of relevant SNOMED CT codes recorded in GP records.
Using the GP Connect: Update Record technical documentation [@NHSEngland2024b] and OpenCodelists (https://www.opencodelists.org/), we identified three codes associated with Pharmacy First consultations (see Table A1 in Supplementary Materials).

### Pharmacy First clinical conditions

The Pharmacy First service includes the supply of medication for the following seven clinical conditions: acute otitis media, impetigo, infected insect bites, shingles, sinusitis, sore throat, and uncomplicated urinary tract infections [@NHSEngland2024].
To identify these clinical conditions within Pharmacy First consultations, we implemented the clinical pathways codelist.
This codelist included a single code for each condition, with each code being specifically used for GP Connect: Update Record. 
Each clinical condition has its own specific inclusion and exclusion criteria [@NHSEngland2023c].
In this study, all individuals were included without applying any specific inclusion or exclusion criteria. 

### Pharmacy First medications

We created codelists for each of the seven clinical conditions based on the medications specified in the Pharmacy First Patient Group Directions [@NHSEngland2023], see Table A1 in Supplementary Materials.

## Software and Reproducibility
Data management was performed using Python v3.9.1 and R v4.0.5.
Code for data management and analysis, as well as codelists, are available for review and reuse under the MIT open license at https://github.com/opensafely/pharmacy-first/.

# Results

## Cohort descriptions

```{r, message=FALSE, warning=FALSE}
population_table <- read_csv(here("released_output", "population", "pf_tables.csv"))
View(population_table)
df_demographics_table <- population_table %>% 
  filter(table=="tab_demographics") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  select(-table) %>%
  mutate(category = factor(category, levels = c("sex", "age_band", "ethnicity", "imd", "region"), 
  labels = c("Sex", "Age Band", "Ethnicity", "IMD", "Region"))) %>%
  mutate(subcategory = recode(subcategory, 
    "female" = "Female", 
    "male" = "Male"
  ))

# Addressing an ordering problem (see PR description)
df_demographics_table <- df_demographics_table %>%  
  mutate(subcategory = if_else(
    category == "Ethnicity", 
    factor(subcategory, levels = ethnicity_levels, labels = ethnicity_labels), 
    factor(subcategory, levels = subcategory_levels, labels = subcategory_labels)
  )) %>%
  arrange(category, match(subcategory, ethnicity_labels)) %>% 
  group_by(category) %>% 
  arrange(category, 
          ifelse(subcategory == "Missing", Inf, as.numeric(subcategory))) %>%
  ungroup()


gt_demographics(df_demographics_table)

```
**Table 1**. Cohort description for all patients with at least one Pharmacy First consultation code during 12 months following the services’ launch date 31 January 2024.

## Count of Pharmacy First Consultations

### Total

```{r plot-pf-consultations, message=FALSE, warning=FALSE, fig.height=8, fig.width=10, fig.cap="**Figure 1**. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP."}
# Create figure for total count of Pharmacy First consultations for each code (3 codes)
df_pf_consultations <- df_measures |>
  filter(measure_desc == "clinical_service") |>
  filter(is.na(group_by)) |>
  select(measure, interval_start, numerator) |>
  mutate(measure = factor(measure,
    levels = c(
      "Consultation Service",
      "Pharmacy First Consultation",
      "Community Pharmacy First Service"
    ),
    labels = c(
      "CP Consultation Service for minor illness (1577041000000109)",
      "Pharmacy First service (983341000000102)",
      "CP Pharmacy First Service (2129921000000100)"
    )
  )) |>
  group_by(interval_start) |>
  mutate(
    pf_consultation_total = sum(numerator, na.rm = TRUE),
    data_desc = "Pharmacy First Consultation"
  ) |>
  rename(pf_code = measure) |>
  filter(interval_start >= "2024-02-01") |>
  ungroup() |>
  pivot_longer(
    cols = c(numerator, pf_consultation_total),
    names_to = "measure"
  ) |>
  mutate(
    measure = case_when(
      measure == "pf_consultation_total" ~ "Pharmacy First Service Total",
      measure == "numerator" ~ pf_code
    ),
    data_desc = case_when(
      measure == "Pharmacy First Service Total" ~ "total",
      measure %in% c(
        "CP Consultation Service for minor illness (1577041000000109)",
        "Pharmacy First service (983341000000102)",
        "CP Pharmacy First Service (2129921000000100)"
      ) ~ "breakdown"
    ),
    measure = factor(measure,
      levels = c(
        "Pharmacy First Service Total",
        "CP Consultation Service for minor illness (1577041000000109)",
        "Pharmacy First service (983341000000102)",
        "CP Pharmacy First Service (2129921000000100)"
      )
    ),
    data_desc = factor(data_desc,
      levels = c("total", "breakdown"),
      labels = c("Total", "Breakdown")
    )
  ) |>
  select(measure, interval_start, data_desc, value)

plot_pf_consultations <- df_pf_consultations |>
  ggplot(
    aes(
      x = interval_start,
      y = value,
      colour = measure,
      shape = measure,
      fill = measure
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(alpha = .2) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  guides(
    color = guide_legend(nrow = 2),
    shape = guide_legend(nrow = 2)
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = NULL,
    colour = NULL,
    shape = NULL,
    fill = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14)
  ) +
  facet_wrap(
    ~data_desc,
    ncol = 1
  ) +
  scale_colour_viridis_d(option = "plasma", end = .75) +
  scale_y_continuous(
    labels = scales::number,
    breaks = c(seq(0, 60000, by = 10000))
  )

plot_pf_consultations
```

### Demographic breakdowns

#### Sex

```{r plot-pf-consultations-by-sex, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="**Figure 2**. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by sex."}
df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Sex") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = sex,
    guide_nrow = 1,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  )
```

#### Age band

```{r plot-pf-consultations-by-age-band, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="**Figure 3**. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by age bands."}
unique_categories <- sort(unique(df_measures$age_band))
manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 1)
shape_mapping <- setNames(manual_ggplot2_shapes[1:length(unique_categories)], unique_categories)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Age band") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = age_band,
    guide_nrow = 1,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping)
```

#### Index of Multiple Deprivation (IMD)

```{r plot-pf-consultations-by-imd, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="**Figure 4**. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by IMD."}
unique_categories <- sort(unique(df_measures$imd))

manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 1)
shape_mapping <- setNames(
  manual_ggplot2_shapes[1:length(unique_categories)],
  unique_categories
)

# Get one more colour than I need from viridis so I avoid using yellow in the plot
# paletteer_c("viridis::viridis", n = 7)
manual_ggplot2_colours <- c(
  "#440154FF",
  "#443A83FF",
  "#31688EFF",
  "#21908CFF",
  "#35B779FF",
  "#8FD744FF",
  "#FDE725FF"
)
colour_mapping <- setNames(
  c(manual_ggplot2_colours[1:(length(unique_categories) - 1)], "gray40"),
  unique_categories
)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "IMD") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = imd,
    guide_nrow = 1,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping) +
  scale_colour_manual(values = colour_mapping) +
  scale_fill_manual(values = colour_mapping)
```

#### Region

```{r plot-pf-consultations-by-region, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="**Figure 5**. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by region."}
unique_categories <- sort(unique(df_measures$region))

manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 8, 4, 10, 7, 1)
shape_mapping <- setNames(
  manual_ggplot2_shapes[1:length(unique_categories)],
  unique_categories
)

# Get one more colour than I need from viridis so I avoid using yellow in the plot
# paletteer::paletteer_c("viridis::viridis", n = 11)
manual_ggplot2_colours <- c(
  "#440154FF",
  "#482576FF",
  "#414487FF",
  "#35608DFF",
  "#2A788EFF",
  "#21908CFF",
  "#22A884FF",
  "#43BF71FF",
  "#7AD151FF",
  "#BBDF27FF",
  "#FDE725FF"
)

colour_mapping <- setNames(
  c(manual_ggplot2_colours[1:(length(unique_categories) - 1)], "gray40"),
  unique_categories
)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Region") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = region,
    guide_nrow = 2,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping) +
  scale_colour_manual(values = colour_mapping) +
  scale_fill_manual(values = colour_mapping)
```

#### Ethnicity

```{r plot-pf-consultations-by-ethnicity, message=FALSE, warning=FALSE, fig.height=15, fig.width=10, fig.cap="**Figure 5**. Monthly count of patients with a Pharmacy First consultation code in OpenSAFELY-TPP broken down by region."}
unique_categories <- sort(unique(df_measures$ethnicity))

manual_ggplot2_shapes <- c(15, 19, 24, 23, 3, 1)
shape_mapping <- setNames(
  manual_ggplot2_shapes[1:length(unique_categories)],
  unique_categories
)

# Get one more colour than I need from viridis so I avoid using yellow in the plot
# paletteer::paletteer_c("viridis::viridis", n = 6)
manual_ggplot2_colours <- c(
  "#440154FF",
  "#414487FF",
  "#2A788EFF",
  "#22A884FF",
  "#7AD151FF",
  "#FDE725FF"
)

colour_mapping <- setNames(
  c(manual_ggplot2_colours[1:(length(unique_categories) - 1)], "gray40"),
  unique_categories
)

df_measures %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) |>
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Ethnicity") |>
  plot_measures(
    select_value = numerator,
    select_interval_date = interval_start,
    colour_var = ethnicity,
    guide_nrow = 2,
    facet_wrap = FALSE,
    add_vline = FALSE,
    title = NULL,
    y_label = NULL
  ) +
  scale_shape_manual(values = shape_mapping) +
  scale_colour_manual(values = colour_mapping) +
  scale_fill_manual(values = colour_mapping) +
  facet_wrap(~ethnicity, ncol = 1, scales = "free_y")
```

## Linkage of Pharmacy First Consultations

```{r plot-pf-med-condition-linkage, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="**Figure 6**. Monthly breakdown of Pharmacy First consultation codes linked to a Pharmacy First medication, clinical condition, or both in OpenSAFELY-TPP."}
df_pf_descriptive_stats <- df_descriptive_stats %>%
  filter(measure %in% c("pfmed_with_pfid", "pfcondition_with_pfid", "pfmed_and_pfcondition_with_pfid")) %>%
  mutate(
    measure = factor(measure,
      levels = c(
        "pfmed_with_pfid",
        "pfcondition_with_pfid",
        "pfmed_and_pfcondition_with_pfid"
      ),
      labels = c(
        "Medication",
        "Clinical condition",
        "Both"
      )
    )
  ) %>%
  left_join(df_pf_consultations_total,
    by = c("interval_start")
  )

df_pf_descriptive_stats <- df_pf_descriptive_stats %>%
  filter(between(interval_start, as.Date("2024-02-01"), as.Date("2025-02-01"))) %>%
  group_by(interval_start) %>%
  mutate(measure = factor(measure, levels = c("Both", "Clinical condition", "Medication"))) |>
  arrange(desc(measure), .by_group = TRUE) |>
  mutate(
    numerator_exclusive = case_when(
      measure == "Both" ~ numerator,
      measure %in% c("Clinical condition", "Medication") ~ numerator - numerator[measure == "Both"]
    ),
    ratio_exclusive = numerator_exclusive / pf_consultation_total,
    cumulative_ratio_exclusive = case_when(interval_start == max(df_pf_descriptive_stats$interval_start) ~ cumsum(ratio_exclusive), TRUE ~ NA)
  )

plot_pf_med_condition_linkage <- df_pf_descriptive_stats |>
  ggplot(
    aes(
      x = interval_start,
      y = ratio_exclusive,
      fill = measure
    )
  ) +
  geom_area(
    alpha = 0.85,
    linewidth = .5,
    colour = "white"
  ) +
  geom_label(
    aes(
      y = cumulative_ratio_exclusive,
      label = ifelse(is.na(ratio_exclusive), "", scales::percent(ratio_exclusive, accuracy = .1)),
    ),
    vjust = -0.05,
    size = 3.5,
    show.legend = FALSE,
    color = "black",
    fill = "white"
  ) +
  scale_fill_viridis_d() +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Pharmacy First consultations linked to: "
  ) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14)
  )

plot_pf_med_condition_linkage
```

### Pharmacy First Clinical Pathways

```{r, message=FALSE, warning=FALSE}
df_clinical_pathways_by_sex <- population_table %>% 
  filter(table=="tab_pf_pathways_by_sex") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  select(-table) %>% 
  pivot_wider(names_from = category, values_from = numerator) %>% 
  group_by(subcategory) %>% 
  mutate(total = female + male) %>% 
  ungroup() %>% 
  mutate(pct = total/sum(total) * 100,
  subcategory = factor(subcategory, levels = c(
    "otitismedia",
    "impetigo",
    "insectbite",
    "shingles",
    "sinusitis",
    "sorethroat",
    "uti"),
  labels = c(
    "Acute otitis media", 
    "Impetigo", 
    "Infected insect bites", 
    "Shingles", 
    "Sinusitis", 
    "Sore throat", 
    "Uncomplicated UTI")))

gt_pathways(df_clinical_pathways_by_sex)
```

- Add existing breakdown figures for: Age band, Ethnicity, IMD (or table to start with)

### Pharmacy First Medications

```{r, message=FALSE, warning=FALSE}
df_consultation_med_counts <- read_csv(
  here("released_output", "measures", "pf_medications_measures_tidy.csv"),
  col_types = cols(dmd_code = col_character())
)
vmp_lookup <- read_csv(
  here("lib", "reference", "vmp_vtm_lookup.csv"),
  col_types = cols(id = col_character())
) %>%
  rename(code = id)

df_pf_med_counts <- df_consultation_med_counts |>
  select(numerator, code = dmd_code, pharmacy_first_med) |>
  left_join(vmp_lookup, by = "code") |>
  filter(numerator > 0) |>
  select(-code) %>%
  group_by(pharmacy_first_med, vmp_nm) |>
  summarise(count = sum(numerator, na.rm = TRUE)) |>
  filter(!is.na(vmp_nm)) %>%
  ungroup() |>
  group_by(pharmacy_first_med) |>
  mutate(ratio_by_group = count / sum(count, na.na.rm = TRUE)) |>
  slice_max(order_by = ratio_by_group, n = 10) |>
  ungroup()

df_pf_and_non_pf_med_counts <- df_pf_med_counts %>%
  arrange(!pharmacy_first_med) %>%
  mutate(pharmacy_first_med = factor(pharmacy_first_med,
    levels = c(FALSE, TRUE),
    labels = c(("Medication not included in codelists"), "Medication included in codelists")
  )) %>%
  group_by(pharmacy_first_med)

gt_top_meds(df_pf_and_non_pf_med_counts)
```
**Table 3**. Top ten medications linked to Pharmacy First consultations during the first 12 months following the service's launch date on 31 January 2024.

## Comparison with NHS BSA data

### Pharmacy First Consultations

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.cap="**Figure 7**. Monthly counts and ratio of patients with a Pharmacy First consultation code in OpenSAFELY-TPP and the NHS BSA comparison dataset."}
# Ideally I would like to load this at the top of this script
# But this file uses datasets created further down so it easiest to load it here
df_pf_consultations_total <- df_pf_consultations |>
  filter(data_desc == "Total") |>
  mutate(
    data_desc = "Pharmacy First Consultation"
  ) |>
  select(interval_start, pf_consultation_total = value, data_desc) |>
  unique()

source(here("lib", "functions", "combine_os_nhsbsa_validation_data.R"))

# Pharmacy First Consultations
# Instead of only using the count of PF consultations in OS with a clinical
# condition we are using all PF consultations with a consultation id
df_pf_consultation_validatio_plot <- pf_validation |>
  filter(
    (data_source == "opensafely" &
      data_desc == "pf_consultation" &
      count_desc == "consultation_id") |
      (data_source == "nhs_bsa" &
        data_desc == "pf_consultation" &
        count_desc == "consultation_type")
  )

# Pivot wider to calculate ratio
df_pf_consultation_validatio_plot <- df_pf_consultation_validatio_plot |>
  select(date, data_source, count_method, count_total) |>
  pivot_wider(names_from = c(data_source, count_method), values_from = count_total) |>
  mutate(ratio = opensafely_opensafely_tpp / nhs_bsa_count_100pct)

# Select colours for groups
# Group 1:
colour_group_1 <- "#CF4446FF"
# Group 1: #1B0C42FFpaletteer_c("viridis::inferno", n = 10)[2]
colour_group_2 <- "#1B0C42FF"

# Create consultation validation plot
pf_validatio_consultation_plot <- df_pf_consultation_validatio_plot |>
  ggplot(aes(x = date)) +
  geom_point(
    aes(
      y = nhs_bsa_count_100pct,
      colour = "NHS BSA",
      shape = "NHS BSA"
    ),
    size = 2.6
  ) +
  geom_line(
    aes(y = nhs_bsa_count_100pct, colour = "NHS BSA"),
    alpha = .3,
    linewidth = .5
  ) +
  geom_point(
    aes(
      y = opensafely_opensafely_tpp,
      colour = "OpenSAFELY-TPP",
      shape = "OpenSAFELY-TPP"
    ),
    size = 2.6
  ) +
  geom_line(
    aes(
      y = opensafely_opensafely_tpp,
      colour = "OpenSAFELY-TPP"
    ),
    alpha = .3,
    linewidth = .5
  ) +
  geom_label(
    aes(
      y = opensafely_opensafely_tpp,
      label = scales::percent(ratio, accuracy = .1),
    ),
    colour = colour_group_2,
    vjust = -.5,
    size = 3.5
  ) +
  scale_y_continuous(labels = scales::number) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  scale_colour_manual(
    values = c(
      "NHS BSA" = colour_group_1,
      "OpenSAFELY-TPP" = colour_group_2
    ),
    name = NULL
  ) +
  scale_shape_manual(
    values = c(
      "NHS BSA" = 15,
      "OpenSAFELY-TPP" = 16
    ),
    name = NULL
  ) +
  labs(
    x = NULL,
    y = NULL,
    colour = NULL,
    shape = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14)
  )

pf_validatio_consultation_plot
```


### Pharmacy First Medications


# References
