---
title: "Pharmacy First"
output:
  html_document:
    toc: true
    toc_depth: 4
  pdf_document: default
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: references.bib
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(readr)
library(gt)
library(patchwork)
```

```{r load-data, message=FALSE, warning=FALSE}
# Load functions
source(here("lib", "functions", "tidy_measures.R"))
source(here("lib", "functions", "plot_measures.R"))

# Load validation data:
# - df_bsa_medication_validation: date, pharmacy_advanced_service, bnf_paragraph, count
# - df_bsa_consultation_validation: date, consultation_type, source, count_method, count
source(here("lib", "functions", "load_validation_data.R"))

# Load opensafely ouputs:
# - df_measures: measure, interval_start, interval_end, ratio numerator, denominator, age_band, sex,imd, region, ethnicity
# - df_descriptive_stats: measure, interval_start, interval_end, ratio numerator, denominator
# - df_pfmed: measure, interval_start, interval_end, ratio, numerator, denominator, dmd_code
# - df_condition_provider: measure, interval_start, interval_end, ratio, numerator, denominator, pf_status, imd
source(here("lib", "functions", "load_opensafely_outputs.R"))
```

#### Pregnancy Codelist

The [Pregnancy Codelist](https://www.opencodelists.org/codelist/nhsd-primary-care-domain-refsets/preg_cod/20200812/#full-list) was used to identify patients who were pregnant during each month.

#### Ethnicity Codelist

We used the [Ethnicity Codelist](https://www.opencodelists.org/codelist/opensafely/ethnicity-snomed-0removed/2e641f61/) identify ethnicity in Electronic Health Records.
To ensure comprehensive ethnicity data, we supplemented missing ethnicity values with data from the Secondary Uses Service (SUS).

# Results

### Total population

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=10}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(is.na(group_by)) |>
  select(measure, interval_start, numerator) |>
  mutate(measure = factor(measure,
    levels = c("Consultation Service", "Pharmacy First Consultation"),
    labels = c(
      "Consultation Service for minor illness (1577041000000109)",
      "Pharmacy First service (983341000000102)"
    )
  ))

fig_pf_individual_consultations_count <- df_measures_selected |>
  select(measure, interval_start, numerator) |>
  ggplot(aes(
    x = interval_start,
    y = numerator,
    colour = measure,
    shape = measure,
  )) +
  geom_point(size = 2) +
  geom_line(alpha = .3) +
  labs(
    title = NULL,
    x = NULL,
    y = "Total count",
    colour = NULL,
    shape = NULL
  ) +
  scale_y_continuous(
    labels = scales::label_number(),
  ) +
  theme(legend.position = "bottom") +
  guides(
    colour = guide_legend(ncol = 2),
    shape = guide_legend(ncol = 2)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  geom_vline(
    xintercept = lubridate::as_date(c(
      "2024-01-31"
    )),
    linetype = "dotted",
    colour = "orange",
    size = .7
  ) +
  scale_colour_viridis_d(end = .75) +
  theme(
    text = element_text(size = 14)
  )

ggsave(
  here("released_output", "results", "figures", "fig_pf_individual_consultations_count.png"),
  fig_pf_individual_consultations_count,
  width = 10, height = 6
)

fig_pf_individual_consultations_count

fig_pf_grouped_consultations_count <- df_measures_selected |>
  group_by(interval_start) |>
  mutate(
    pf_consultation_total = sum(numerator, na.rm = TRUE),
    data_desc = "Pharmacy First Consultation"
  ) |>
  ggplot(aes(
    x = interval_start,
    y = pf_consultation_total,
    colour = data_desc,
    shape = data_desc,
  )) +
  geom_point(size = 2) +
  geom_line(alpha = .3) +
  labs(
    title = NULL,
    x = NULL,
    y = "Total count",
    colour = NULL,
    shape = NULL
  ) +
  scale_y_continuous(
    labels = scales::label_number(),
  ) +
  theme(legend.position = "bottom") +
  guides(
    colour = guide_legend(ncol = 2),
    shape = guide_legend(ncol = 2)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    labels = scales::label_date_short()
  ) +
  geom_vline(
    xintercept = lubridate::as_date(c(
      "2024-01-31"
    )),
    linetype = "dotted",
    colour = "orange",
    size = .7
  ) +
  scale_colour_viridis_d(end = .75) +
  theme(
    text = element_text(size = 14)
  )

ggsave(
  here("released_output", "results", "figures", "fig_pf_grouped_consultations_count.png"),
  fig_pf_grouped_consultations_count,
  width = 10, height = 6
)
```

```{r, message=FALSE, warning=FALSE, fig.height=10, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(is.na(group_by))

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions",
)
```

### Breakdown by age

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)
```

### Breakdown by sex

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = sex_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = sex_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = sex_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = sex_palette)
```

### Breakdown by IMD

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = gradient_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = gradient_palette)
```

### Breakdown by region

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = region_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = region_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = region_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = region_palette)
```

### Breakdown by ethnicity

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Consultations",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = ethnicity_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "pharmacy_first_services") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Consultations per 1000 people",
  y_label = "Number of codes for FP consultations",
) + scale_color_manual(values = ethnicity_palette)
```

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Pharmacy First Conditions",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = ethnicity_palette)

# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Ethnicity")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = ratio,
  select_interval_date = interval_end,
  colour_var = ethnicity,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
  title = "Rate of Pharmacy First Conditions per 1000 people",
  y_label = "Number of codes for PF conditions"
) + scale_color_manual(values = ethnicity_palette)
```
```{r, message=FALSE, warning=FALSE, echo = FALSE}
# OpenSAFELY data for clinical conditions into a tidy df
df_opensafely_validation <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  # filter(interval_start >= as.Date("2024-02-01") & interval_start <= as.Date("2024-07-30")) %>%
  filter(is.na(group_by)) %>%
  select(date = interval_start, consultation_type = measure, count = numerator) %>%
  mutate(
    source = "opensafely",
    count_method = "opensafely_tpp"
  ) |>
  filter(date >= "2024-01-01") %>%
  relocate(date, consultation_type, source, count_method, count)

# Combining rows from OS and BSA validation dataframes
df_validation_condition_counts <- bind_rows(df_opensafely_validation, df_bsa_consultation_validation)

# Line graph comparing clinical condition counts of BSA and OS data
fig_validation_condition_count <- df_validation_condition_counts %>%
  filter(count_method %in% c("opensafely_tpp", "count_40pct")) %>%
  mutate(source = factor(source,
    levels = c("opensafely", "nhs_bsa"),
    labels = c("OpenSAFELY-TPP", "NHS BSA (40%)")
  )) %>%
  ggplot(
    aes(
      x = date,
      y = count,
      shape = consultation_type,
      color = consultation_type,
      fill = consultation_type,
      group = consultation_type
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  facet_wrap(~source, scales = "free_y") +
  scale_x_date(
    labels = scales::label_date_short()
  ) +
  labs(x = NULL, y = "Count", colour = NULL, shape = NULL, fill = NULL) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_shape_manual(
    values = c(
      "Acute Sinusitis" = 15,
      "Infected Insect Bite" = 19,
      "UTI" = 4,
      "Acute Otitis Media" = 23,
      "Acute Pharyngitis" = 3,
      "Herpes Zoster" = 17,
      "Impetigo" = 8
    )
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_y_continuous(labels = scales::number)

# Another plot visualising the percentage
fig_validation_condition_pct <- df_validation_condition_counts %>%
  filter(count_method %in% c("opensafely_tpp", "count_40pct")) %>%
  pivot_wider(names_from = c(source, count_method), values_from = count) %>%
  mutate(source = "Percentage of NHS BSA (40%) in OpenSAFELY") %>%
  ggplot(
    aes(
      x = date,
      y = opensafely_opensafely_tpp / nhs_bsa_count_40pct,
      shape = consultation_type,
      color = consultation_type,
      fill = consultation_type,
      group = consultation_type
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  facet_wrap(~source, scales = "free_y") +
  scale_x_date(
    labels = scales::label_date_short()
  ) +
  labs(x = NULL, y = "Percent", colour = NULL, shape = NULL, fill = NULL) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_shape_manual(
    values = c(
      "Acute Sinusitis" = 15,
      "Infected Insect Bite" = 19,
      "UTI" = 4,
      "Acute Otitis Media" = 23,
      "Acute Pharyngitis" = 3,
      "Herpes Zoster" = 17,
      "Impetigo" = 8
    )
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_y_continuous(labels = scales::percent)

# Combining the plots with patchwork
fig_validation_condition_count_pct <- (fig_validation_condition_count + fig_validation_condition_pct) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect", widths = c(2, 1)) &
  theme(
    legend.position = "bottom",
    text = element_text(size = 15),
    strip.background = element_rect(size = 0),
    strip.text.x = element_text(size = 13, face = "bold")
  )

fig_validation_condition_count_pct

ggsave(
  here("released_output", "results", "figures", "fig_validation_condition_count_pct.png"),
  fig_validation_condition_count_pct,
  width = 15, height = 6
)
```

```{r, message=FALSE, warning=FALSE, echo = FALSE, fig.width=8}
# Line graph comparing clinical condition counts of BSA and OS data
fig_pf_descriptive_stats <- df_descriptive_stats %>%
  mutate(
    measure = factor(measure,
      levels = c("pf_with_pfmed", "pf_with_pfcondition", "pf_with_pfmed_and_pfcondition"),
      labels = c("PF Med", "PF Condition", "PF Med & PF Condition")
    )
  ) |>
  ggplot(aes(
    x = interval_start,
    y = ratio,
    colour = measure,
    shape = measure,
  )) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  labs(
    x = NULL,
    y = NULL,
    shape = "PF consultation linked to:",
    colour = "PF consultation linked to:"
  ) +
  scale_x_date(
    labels = scales::label_date_short(), breaks = "month"
  ) +
  scale_y_continuous(
    labels = scales::percent,
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_colour_brewer(palette = "Dark2")


fig_pf_descriptive_stats

ggsave(
  here("released_output", "results", "figures", "fig_pf_descriptive_stats.png"),
  fig_pf_descriptive_stats,
  width = 10, height = 6
)
```

```{r, message=FALSE, warning=FALSE, echo = FALSE}
# Validation of pharmacy first medication counts figure
# OS data - waiting on released output

df_bsa_medication_validation_sum <- df_bsa_medication_validation %>%
  group_by(date) %>%
  summarise(count = sum(count) * 0.4) %>%
  mutate(
    source = "nhs_bsa",
    count_method = "count_40pct"
  )
range(df_pfmed$interval_start)
df_opensafely_pfmed_sum <- df_pfmed %>%
  rename(date = interval_start) %>%
  group_by(date) %>%
  summarise(count = sum(numerator)) %>%
  mutate(
    source = "opensafely_tpp",
    count_method = "opensafely_tpp"
  )

df_validation_med_counts <- bind_rows(df_opensafely_pfmed_sum, df_bsa_medication_validation_sum) |>
  filter(date >= "2024-01-01" & date <= "2024-07-01")

fig_validation_med_count <- df_validation_med_counts |>
  mutate(
    source = factor(source, levels = c("opensafely_tpp", "nhs_bsa"), labels = c("OpenSAFELY-TPP", "NHS BSA")),
    count_method = factor(count_method, levels = c("opensafely_tpp", "count_40pct"), labels = c("OpenSAFELY-TPP", "NHS BSA (40%)"))
  ) |>
  ggplot(aes(
    x = date,
    y = count,
    colour = count_method,
    shape = count_method
  )) +
  geom_point(size = 2) +
  facet_wrap(~count_method, scales = "free_y") +
  geom_line(size = 0.5) +
  labs(
    x = NULL,
    y = "Count",
    colour = NULL,
    shape = NULL,
  ) +
  scale_x_date(
    labels = scales::label_date_short(), breaks = "month"
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_colour_brewer(palette = "Dark2")

# Another plot visualising the percentage
fig_validation_med_pct <- df_validation_med_counts %>%
  filter(count_method %in% c("opensafely_tpp", "count_40pct")) %>%
  pivot_wider(names_from = c(source, count_method), values_from = count) %>%
  mutate(source = "Percentage of NHS BSA (40%) in OpenSAFELY") %>%
  ggplot(
    aes(
      x = date,
      y = opensafely_tpp_opensafely_tpp / nhs_bsa_count_40pct,
      shape = source,
      color = source,
      fill = source,
      group = source
    )
  ) +
  geom_point(size = 2.5) +
  geom_line(size = 0.5) +
  facet_wrap(~source, scales = "free_y") +
  scale_x_date(
    labels = scales::label_date_short(), breaks = "month"
  ) +
  labs(x = NULL, y = "Percent", colour = NULL, shape = NULL, fill = NULL) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.9
  ) +
  scale_shape_manual(
    values = c("Percentage of NHS BSA (40%) in OpenSAFELY" = 15)
  ) +
  theme(
    text = element_text(size = 14)
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_y_continuous(labels = scales::percent)

# Combining the plots with patchwork
fig_validation_medication_count_pct <- (fig_validation_med_count + fig_validation_med_pct) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect", widths = c(2, 1)) &
  theme(
    legend.position = "bottom",
    text = element_text(size = 15),
    strip.background = element_rect(size = 0),
    strip.text.x = element_text(size = 13, face = "bold")
  )

fig_validation_medication_count_pct

ggsave(
  here("released_output", "results", "figures", "fig_validation_medication_count_pct.png"),
  fig_validation_medication_count_pct,
  width = 15, height = 6
)
```
```{r, message=FALSE, warning=FALSE, echo = FALSE, fig.width=8}
# GP vs PF provider graph

df_condition_provider_grouped <- df_condition_provider %>%
  group_by(measure, interval_start, pf_status) %>%
  summarise(
    count = sum(numerator)
  ) %>%
  mutate(
    measure = factor(measure,
      levels = c(
        "count_acute_sinusitis_total",
        "count_infected_insect_bite_total",
        "count_uncomplicated_urinary_tract_infection_total",
        "count_acute_otitis_media_total",
        "count_acute_pharyngitis_total",
        "count_herpes_zoster_total",
        "count_impetigo_total"
      ),
      labels = c(
        "Acute Sinusitis",
        "Infected Insect Bite",
        "UTI",
        "Acute Otitis Media",
        "Acute Pharyngitis",
        "Herpes Zoster",
        "Impetigo"
      )
    ),
    pf_status = factor(pf_status,
      levels = c(TRUE, FALSE),
      labels = c("Linked to Pharmacy First consultation", "Not linked to Pharmacy First consultation")
    )
  )

fig_pf_condition_provider_count <- ggplot(
  df_condition_provider_grouped,
  aes(
    x = interval_start,
    y = count,
    colour = pf_status,
    shape = pf_status
  )
) +
  geom_point(size = 1.5) +
  geom_line(size = 0.5) +
  facet_wrap(~measure, scales = "free_y") +
  labs(
    x = NULL, y = "Count", color = NULL, shape = NULL
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    axis.title.x = element_blank()
  ) +
  scale_x_date(
    labels = scales::label_date_short()
  ) +
  geom_vline(
    xintercept = lubridate::as_date("2024-02-01"),
    linetype = "dotted",
    colour = "orange",
    linewidth = .7
  ) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.75
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 14),
    strip.background = element_rect(size = 0),
    # strip.text.x = element_text(size = 13, face = "bold")
  )

fig_pf_condition_provider_count

ggsave(
  here("released_output", "results", "figures", "fig_pf_condition_provider_count.png"),
  fig_pf_condition_provider_count,
  width = 13, height = 8
)
```

# References
