---
title: "Pharmacy First"
output:
  html_document:
    toc: true
    toc_depth: 4
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(readr)
```

```{r load-data, message=FALSE, warning=FALSE}
# Load plotting function
source(here::here("lib", "functions", "function_tidy_measures.R"))
source(here::here("lib", "functions", "function_plot_measures.R"))

# Load data
df_measures <- readr::read_csv(
  here::here("output", "measures", "pf_codes_conditions_measures.csv")
)

# Define dictionaries with tidy names and mappings for measures
pf_measures_name_dict <- list(
  blood_pressure_service = "Blood Pressure Service",
  contraception_service = "Contraception Service",
  consultation_service = "Consultation Service",
  pharmacy_first_service = "Pharmacy First Service",
  acute_otitis_media = "Acute Otitis Media",
  herpes_zoster = "Herpes Zoster",
  acute_sinusitis = "Acute Sinusitis",
  impetigo = "Impetigo",
  infected_insect_bite = "Infected Insect Bite",
  acute_pharyngitis = "Acute Pharyngitis",
  uncomplicated_urinary_tract_infection = "UTI"
)

pf_measures_name_mapping <- list(
  blood_pressure_service = "clinical_service",
  contraception_service = "clinical_service",
  consultation_service = "clinical_service",
  pharmacy_first_service = "clinical_service",
  acute_otitis_media = "clinical_condition",
  herpes_zoster = "clinical_condition",
  acute_sinusitis = "clinical_condition",
  impetigo = "clinical_condition",
  infected_insect_bite = "clinical_condition",
  acute_pharyngitis = "clinical_condition",
  uncomplicated_urinary_tract_infection = "clinical_condition"
)

pf_measures_groupby_dict <- list(
  age_band = "Age band",
  sex = "Sex",
  imd = "IMD",
  region = "Region",
  ethnicity = "Ethnicity"
)

df_measures <- tidy_measures(
  data = df_measures,
  pf_measures_name_dict = pf_measures_name_dict,
  pf_measures_name_mapping = pf_measures_name_mapping,
  pf_measures_groupby_dict = pf_measures_groupby_dict
)
```

# Background

Add background here.

# Methods

This study used data from OpenSAFELY-TPP, which covers 40% of the population of England. For a description of the representativeness of this sample, please see our manuscript [here](https://wellcomeopenresearch.org/articles/7-191/v1).
Individuals were included if they were alive and registered at a TPP practice each month, across the study period.
Patients were excluded if their listed age was not between 0 and 120 years.
Counts represent patients with at least one clinical code of relevance in that month.
Patients with more than one of the same clinical code in a month were only counted once. Rates divide the count by the included study population and multiply by 1,000 to achieve a rate per 1,000 registered patients.
Counts <=7 have been redacted and all numbers rounded to the nearest 5 to avoid potential re-identification of individuals. The rates displayed were computed with these rounded counts.
Our data relies on a relevant Pharmacy First code being added to a patient's GP record.
The Pharmacy First service relies on [GP Connect - Update Record](https://digital.nhs.uk/services/gp-connect/gp-connect-in-your-organisation/gp-connect-update-record) to update a patient's GP record with consultation information from the community pharmacy.
Following the launch of the Pharmacy First service, there has been a [Gradual roll-out of GP Connect - Update Record](https://cpe.org.uk/our-news/gp-connect-update-record-rollout-and-flow-of-information/) across the approved community pharmacy IT system suppliers.

Links to the codelist for each analysis can be found beneath the relevant section.

# Results

## Clinical Services

### Total population

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=4}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(is.na(group_by))

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = NULL,
  guide_nrow = 1,
  facet_wrap = FALSE,
  facet_var = NULL,
)
```

### Breakdown by age

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure
)
```

### Breakdown by sex

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
)
```

### Breakdown by IMD

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
)
```

### Breakdown by region

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_service") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
)
```

### Clinical Services by ethnicity

```{r, message=FALSE, warning=FALSE}

# TODO

```

## Clinical Pathways

This section focuses on the Clinical Pathways element of the Pharmacy First service.  

## Clinical Condition

This section focuses on the clinical conditions within the Clinical Pathways element of that Pharmacy First service:  
Here we show the number of consultations for each of the Pharmacy First Clinical Pathways Clinical Conditions.

### Total population

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=4}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(is.na(group_by))

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  guide_nrow = 1,
  facet_wrap = FALSE,
  facet_var = NULL,
)
```

### Breakdown by age

```{r, message=FALSE, warning=FALSE, fig.height=15, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Age band")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = age_band,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure
)
```

### Breakdown by sex

```{r, message=FALSE, warning=FALSE, fig.height=15, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Sex")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = sex,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
)
```

### Breakdown by IMD

```{r, message=FALSE, warning=FALSE, fig.height=15, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "IMD")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = imd,
  guide_nrow = 1,
  facet_wrap = TRUE,
  facet_var = measure,
)
```

### Breakdown by region

```{r, message=FALSE, warning=FALSE, fig.height=15, fig.width=8}
# Select measures and breakdown
df_measures_selected <- df_measures %>%
  filter(measure_desc == "clinical_condition") %>%
  filter(group_by == "Region")

# Create visualisation
plot_measures(
  df_measures_selected,
  select_value = numerator,
  select_interval_date = interval_end,
  colour_var = region,
  guide_nrow = 2,
  facet_wrap = TRUE,
  facet_var = measure,
)
```

### Clinical Conditions by ethnicity

```{r, message=FALSE, warning=FALSE}

# TODO

```
